Minimizing post-shock forecasting error using disparate information

### 2020-06-26 manuscript/manuscript.tex
revision to manuscript in discussions and bootstrap section and minor revision to data analysis (and code) in switching to sample estimate of expectations

Things to do:
\begin{enumerate}
  \item Need to add simulations. 
\end{enumerate}

@@

To specify, there are two possible bootstrap procedures for our methods, one based on resampling the donor pool with replacement (\emph{unconditional} on donor pool) and the other without any resampling of the donor pool (\emph{conditional} on donor pool). We denote them by $\mc{B}_u$ and  $\mc{B}_c$, respectively. We first illustrate the simple $\mc{B}_c$. The formal steps of it are outlined in the Supplementary Materials, the intuition for this procedure is as follows: let $B$ be the bootstrap sample size and let $I=\{2, \ldots, n +1\}$ be the indexes for donor pool. Initialize $y_{i,0}$ for all $i \in I$. At iteration $b$, resample the residuals and then obtain shock-effect estimators for each of the disparate time series for all $i \in I$. Then construct and store any of the adjustment estimators $\hat{\alpha}^{(b)}_{\mrm{adj}}$, $\hat{\alpha}^{(b)}_{\mrm{wadj}}$, and $\hat{\alpha}^{(b)}_{\mrm{IVW}}$. We can then estimate distributional quantities of our shock-effect estimators with the bootstrap sample of $\hat{\alpha}^{(b)}_{\mrm{adj}}$, $\hat{\alpha}^{(b)}_{\mrm{wadj}}$, and $\hat{\alpha}^{(b)}_{\mrm{IVW}}$, for $b = 1,\ldots,B$.
%The errors $\varepsilon_{i,t}$ and $\varepsilon_{i,t'}$ for $t \in \{1, \ldots, T_i^*\}$ and $t' \in \{T_i^*+1, \ldots, T_i\}$ are iid. 
%since $y_{i,t}$ differs from $y_{i,t'}$ only in the shock-effect $\alpha_i$ conditioned on $\Theta$, and it is not absorbed into $\varepsilon_{i,t}$ and $\varepsilon_{i,t'}$. Therefore, this parametric bootstrap framework will work for us.
One can then use the bootstrapped shock-effects estimated by our residual bootstrap to provide an approximation for parameters involved in the risk-reduction conditions in Propositions \ref{proprisk} and \ref{propriskwadj2}.

$\mc{B}_u$ is the same as above with the addition of first round of sampling $n$ values from $I=\{2, \ldots, n+1\}$ without replacement to generate  $I^{(b)}$. Then, it follows the same step with the one conditioned on the donor pool by replacing $I$ by $I^{(b)}$. Care is needed when using this bootstrap procedure to estimate the variability of $\hat{\alpha}_{\mrm{wadj}}$. We would assume $\mathbf{W}^*$ is non-degenerate (i.e., no zero weights) in the population when using $\mc{B}_u$.  See Section \ref{discussion} for discussions.

@@

Now, we discuss some consistency results for $\mc{B}_c$ conditioned on $\Theta_1$. Let 
$
  \overline{\hat{\alpha}_{\mrm{adj}}} 
    = \frac{1}{B} \sum_{b=1}^B \sum_{i=2}^{n+1}\hat{\alpha}_i^{(b)}
$, 
$
  \overline{\hat{\alpha}_{\mrm{wadj}}} 
    = \frac{1}{B} \sum_{b=1}^B \sum_{i=2}^{n+1}w_i^* \hat{\alpha}_i^{(b)}
$, 
and 
$
  \overline{\hat{\alpha}_{\mrm{IVW}}} 
    = \frac{1}{B} \sum_{b=1}^B \sum_{i=2}^{n+1}\hat{\alpha}_{{\rm IVW}, i}^{(b)}.
$
The almost-sure convergence results for $\hat{\alpha}_{\mrm{adj}}$ and $\hat{\alpha}_{\mrm{wadj}}$ follow naturally, provided that the regularity conditions outlined in \citet{bose1988edgeworth} hold. The proof for the case of $\overline{\hat{\alpha}_{\mrm{wadj}}}$, the sample mean of the bootstrapped weighted adjustment estimator, can be presented as below. As $B \to \infty$,
\begin{align*}
  \overline{\hat{\alpha}_{\mrm{wadj}}}
  = \frac{1}{B} \sum_{b=1}^B \sum_{i=2}^{n+1}w_i^* \hat{\alpha}_i^{(b)}
  = \sum_{i=2}^{n+1}w_i^* \frac{1}{B} \sum_{b=1}^B\hat{\alpha}_i^{(b)}
  \overset{a.s.}{\rightarrow} \sum_{i=2}^{n+1}w_i^* \alpha_i 
  =\E{\hat{\alpha}_{\rm wadj}|\Theta}.
\end{align*}
The case for the sample variance of the bootstrapped weighted adjustment estimator is similar. The same rationale holds for the adjustment estimator as well since it is a linear combination of OLS estimates, and the weights are not random conditioned on $\Theta$. However, the case for $\hat{\alpha}_{\rm IVW}$ is slightly different because its weights are random. It is not clear with respect to whether similar consistency results holds for $\hat{\alpha}_{\rm IVW}$. We would only claim that our bootstrap procedure provides an approximation for the case of $\hat{\alpha}_{\rm IVW}$.

@@

In particular, $\overline{\hat{\alpha}_{\rm wadj}}$ will provide an approximation for $\E{\alpha_1}$ under $\mc{M}_1$, $\mc{M}_{21}$, and $\mc{M}_{22}$ since it is unbiased for $\E{\alpha_1}$ under those three configurations from Proposition \ref{unbiased}. Therefore, we can estimate the risk-reduction quantities in Section \ref{comparisons} using the estimates from the parametric bootstrap. For example, the $\Delta_{\rm adj}$ and its estimator $\hat{\Delta}_{\rm adj}$ are
\begin{align*}
  \Delta_{\rm adj} 
  &= (\E{\alpha_1})^2 -\var{\hat{\alpha}_{\rm adj}} -(\E{\hat{\alpha}_{\rm adj}}-\E{\alpha_1})^2 \\
  \hat{\Delta}_{\rm adj} 
 & = (\overline{\hat{\alpha}_{\rm wadj}})^2 -S^2_{\hat{\alpha}_{\rm adj}} -(\overline{\hat{\alpha}_{\rm adj}}-\overline{\hat{\alpha}_{\rm wadj}})^2,
\end{align*}
where $S^2_{\hat{\alpha}_{\rm adj}}$ is the bootstrap sample variance for $\hat{\alpha}_{\rm adj}$. Suppose the $\hat{\Delta}_{\rm wadj}$ and $\hat{\Delta}_{\rm IVW}$ are estimators for $\Delta_{\rm wadj}$ and $\Delta_{\rm IVW}$, respectively. Then, we can evaluate the risk-reduction propositions by judging whether $\hat{\Delta}_{\rm est}>0$ for $\text{est} \in \{\text{adj}, \text{wadj}, \text{IVW}\}$, and  the best estimator $\hat{\alpha}_{\rm best}$ is the one with the maximum estimated risk-reduction quantities.

@@

\begin{table}[H]
  \caption{Bootstrap estimates and results yielded by risk-reduction propositions with $B = 1000$}\label{table2}
  \begin{center}
    \begin{tabular}{lrrr}
      & $\hat{\alpha}_{\rm adj}$ & $\hat{\alpha}_{\rm wadj}$ & $\hat{\alpha}_{\rm IVW}$ \\
      \hline 
    Bootstrapped mean  & $-5.340$ & $-5.014$ & $-5.287$ \\
    Bootstrapped variance & 0.508 & 0.480 & 2.821 \\
    \end{tabular}
  \end{center}  
\end{table}
\vspace{-.5cm}

Using the parametric bootstrap proposed in Section \ref{varbootstrap}, we estimated parameters for risk-reduction propositions and risk-reduction quantities proposed in  Section \ref{properties}; and the result is presented in Table \ref{table2}. Plugging these estimates into conditions in Section \ref{properties} and risk-reduction formulas in Section \ref{comparisons} yields (1) $\hat{\alpha}_{\text{adj}}$, $\hat{\alpha}_{\text{wadj}}$, and $\hat{\alpha}_{\text{IVW}}$  reduce the risk and (2) the risk-reduction quantities are $28.031$, $28.084$, and $25.748$, respectively. The estimated risk-reduction quantities  vote $\hat{\alpha}_{\rm wadj}$  for the best estimator. Note that this parametric bootstrap procedure is conditioned  on the  donor pool. Then, we verify the consistency of the result yielded by risk-reduction propositions with the reality as below.

@@

Under the setup of our model, non-uniqueness of $\mathbf{W}^*$ would not be a problem. It is because all risk-reduction propositions and other properties established will still hold. However, non-uniqueness may not be desirable in other model setup. For example, consider the model where $\alpha_i$ is assumed to be not identically distributed, the size of  donor pool to be 2, and there are two solutions to (\ref{SCM}), say, $\mathbf{W}^*_1=(1,0)$ and  $\mathbf{W}^*_2=(0,1)$ with $\var{\alpha_2}\neq \var{\alpha_3}$. In this scenario, the procedure of trying to recover $\alpha_1$ will fail since the resulting $\alpha_1$ will be different with different variances for $\mathbf{W}^*_1=(1,0)$ and  $\mathbf{W}^*_2=(0,1)$. If the non-uniqueness is of concern, users may select the weight that optimizes some objective function. For instance, in the example just discussed, users may select the weight that minimizes the estimated variance of $\hat{\alpha}_{\rm wadj}$.

@@

In terms of the use of our bootstrap procedure, we caution that the bootstrapping residuals in OLS estimation may not provide valid inference in high dimension where $p < T_i$ but $p / T_i$ is not close to zero for $i\in \{2, \ldots, n+1\}$ \citep{el2018can}. Recall that in Section \ref{varbootstrap}, we proposed two possible bootstrap procedures, $\mc{B}_u$  and $\mc{B}_c$. There are some philosophical distinctions between $\mc{B}_u$  and $\mc{B}_c$. $\mc{B}_u$ treats the donor pool as realizations from some infinite super-population of potential donors. In contrast $\mc{B}_c$ treats the donor pool as being fixed  and known before the analysis is conducted, where the randomness comes from parameters and idiosyncratic error. Therefore, it is possible that $\mc{B}_u$ can improve precision of $\mc{B}_2$.

In this study, for the use of $\mc{B}_u$, we assume $\mathbf{W}^*$ is non-degenerate in the population without taking zero weight.  Recall that in Section \ref{constructionofestimators} we note that if there exists some $\mathbf{W}^*$ satisfies (\ref{SCM}) and $2p < n$, there will be infinitely many solutions to $\mathbf{W}^*$. In the simulation setup, we consider $2p >n$ such that the assumption can be reasonably satisfied. However, in the application, it is likely that $\mathbf{W}^*$ may  take values on the boundary of $\mc{W}$, in which case $\mc{B}_u$ may fail for $\hat{\alpha}_{\mrm{wadj}}$ \citep{andrews2000inconsistency}. In this regard, the subsampling in Section 4 of \citet{li2019statistical} may adapt to our scenario to solve this problem by resampling $m$ candidates from donor pool with the conditions $m \to \infty$, and $m/n \to 0$ as $n \to \infty$. But we may leave its mathematical justification for future research. Note that this problem will not occur under  $\mc{B}_c$ because  if $w_i^*=0$ for some $i$ and the weights are known, $w_i^*=0$ makes it impossible for bootstrap estimate of $\hat{\alpha}_i$ to contribute to the bootstrap estimate of distribution of $\hat{\alpha}_{\mrm{wadj}}$. Some recommendations for selection between  $\mc{B}_u$ and $\mc{B}_c$ are as below.

If in practice users detect the boundary case (e.g., $\mathbf{W}^*$ in Section \ref{forecasting}), they are not recommended to use $\mc{B}_u$. Moreover, if $2p < n$, $\mc{B}_u$ is also not recommended since the existence of infinitely many solutions is certain (if there exists some $\mathbf{W}^*$ satisfies (\ref{SCM})), and will guarantee degeneracy of  $\mathbf{W}^*$. Additionally, we  recommend using $\mc{B}_c$ when the donor pool size $n$ is small. When $n$ is large and $2p > n$, the selection would be more up to how users think of the philosophical differences between $\mc{B}_u$ and $\mc{B}_c$ up to application of interest, and the problems that  $\mc{B}_u$ may fail.

>>>>

Things to do:
\begin{enumerate}
  \item Need to add simulations. 
  \item $\mu_{\alpha}=2$, $X_{i,t}\sim \Gamma(1,2)$, $\delta_i, \gamma_i \sim \normal{1}{0.5}$
  \item $\sigma=10$ for the base case.
  \item $\sigma_{\alpha}= (5, 10, 25, 50, 100)$
  \item $\sigma= (5, 10, 25, 50, 100)$
\end{enumerate}

@@

To specify, there are two possible bootstrap procedures for our methods, one based on resampling the donor pool with replacement (\emph{unconditional} on donor pool) and the other without any resampling of the donor pool (\emph{conditional} on donor pool). We denote them by $\mc{B}_u$ and  $\mc{B}_c$, respectively. We first illustrate the simple $\mc{B}_c$. The formal steps of it are outlined in the Supplementary Materials, the intuition for this procedure is as follows: let $B$ be the bootstrap sample size and let $I=\{2, \ldots, n +1\}$ be the indexes for donor pool. Initialize $y_{i,0}$ for all $i \in I$. At iteration $b$, resample the residuals and then obtain shock-effect estimators for each of the disparate time series for all $i \in I$. Then construct and store any of the adjustment estimators $\hat{\alpha}^{(b)}_{\mrm{adj}}$, $\hat{\alpha}^{(b)}_{\mrm{wadj}}$, and $\hat{\alpha}^{(b)}_{\mrm{IVW}}$. We can then estimate distributional quantities of our shock-effect estimators with the bootstrap sample of $\hat{\alpha}^{(b)}_{\mrm{adj}}$, $\hat{\alpha}^{(b)}_{\mrm{wadj}}$, and $\hat{\alpha}^{(b)}_{\mrm{IVW}}$, for $b = 1,\ldots,B$.
%The errors $\varepsilon_{i,t}$ and $\varepsilon_{i,t'}$ for $t \in \{1, \ldots, T_i^*\}$ and $t' \in \{T_i^*+1, \ldots, T_i\}$ are iid. 
%since $y_{i,t}$ differs from $y_{i,t'}$ only in the shock-effect $\alpha_i$ conditioned on $\Theta$, and it is not absorbed into $\varepsilon_{i,t}$ and $\varepsilon_{i,t'}$. Therefore, this parametric bootstrap framework will work for us.
One can then use the bootstrapped shock-effects estimated by our residual bootstrap to provide an approximation for parameters involved in the risk-reduction conditions in Propositions \ref{proprisk} and \ref{propriskwadj2}. $\mc{B}_u$ is the same as above with the addition of first round of sampling $n$ values from $I=\{2, \ldots, n+1\}$ without replacement to generate  $I^{(b)}$. Then, it follows the same step with the one conditioned on the donor pool by replacing $I$ by $I^{(b)}$.

@@

Now, we discuss some consistency results for $\mc{B}_c$ conditioned on $\Theta_1$. The bootstrap is aimed for estimating $\mrm{Var}(\hat{\alpha}_{\rm adj})$, $\mrm{Var}(\hat{\alpha}_{\rm wadj})$, and $\mrm{Var}(\hat{\alpha}_{\rm IVW})$. The consistency results will be similar to the counterparts in sample mean. Therefore, we will briefly discuss the proof for the sample mean case and adapt it to the variance case. Let 
$
  \overline{\hat{\alpha}_{\mrm{adj}}} 
    = \frac{1}{B} \sum_{b=1}^B \sum_{i=2}^{n+1}\hat{\alpha}_i^{(b)}
$, 
$
  \overline{\hat{\alpha}_{\mrm{wadj}}} 
    = \frac{1}{B} \sum_{b=1}^B \sum_{i=2}^{n+1}w_i^* \hat{\alpha}_i^{(b)}
$, 
and 
$
  \overline{\hat{\alpha}_{\mrm{IVW}}} 
    = \frac{1}{B} \sum_{b=1}^B \sum_{i=2}^{n+1}\hat{\alpha}_{{\rm IVW}, i}^{(b)}.
$
The almost-sure convergence results for $\hat{\alpha}_{\mrm{adj}}$ and $\hat{\alpha}_{\mrm{wadj}}$ follow naturally, provided that the regularity conditions outlined in \citet{bose1988edgeworth} hold. The proof for the case of $\overline{\hat{\alpha}_{\mrm{wadj}}}$, the sample mean of the bootstrapped weighted adjustment estimator, can be presented as below. As $B \to \infty$,
\begin{align*}
  \overline{\hat{\alpha}_{\mrm{wadj}}}
  = \frac{1}{B} \sum_{b=1}^B \sum_{i=2}^{n+1}w_i^* \hat{\alpha}_i^{(b)}
  = \sum_{i=2}^{n+1}w_i^* \frac{1}{B} \sum_{b=1}^B\hat{\alpha}_i^{(b)}
  \overset{a.s.}{\rightarrow} \sum_{i=2}^{n+1}w_i^* \alpha_i 
  =\E{\hat{\alpha}_{\rm wadj}|\Theta}.
\end{align*}
The same rationale holds for the adjustment estimator as well since it is a linear combination of OLS estimates, and the weights are not random conditioned on $\Theta$. However, the case for $\hat{\alpha}_{\rm IVW}$ is slightly different because its weights are random. It is not clear with respect to whether similar consistency results holds for $\hat{\alpha}_{\rm IVW}$. We would only claim that our bootstrap procedure provides an approximation for the case of $\hat{\alpha}_{\rm IVW}$.

@@

Recall that  $\hat{\alpha}_{\rm wadj}$ and $\hat{\alpha}_{\rm IVW}$ are unbiased estimators of their expectations, and $\hat{\alpha}_{\rm wadj}$ is an unbiased estimator of $\E{\alpha_1}$ from Proposition \ref{unbiased}. Given the bootstrapped estimates of the variance, we can estimate the risk-reduction quantities in Section \ref{comparisons} using the estimates from the parametric bootstrap. For example, the $\Delta_{\rm adj}$ and its estimator $\hat{\Delta}_{\rm adj}$ are
\begin{align*}
  \Delta_{\rm adj} 
  &= (\E{\alpha_1})^2 -\var{\hat{\alpha}_{\rm adj}} -(\E{\hat{\alpha}_{\rm adj}}-\E{\alpha_1})^2 \\
  \hat{\Delta}_{\rm adj} 
 & = (\hat{\alpha}_{\rm wadj})^2 -S^2_{\hat{\alpha}_{\rm adj}} -(\hat{\alpha}_{\rm adj}-\hat{\alpha}_{\rm wadj})^2,
\end{align*}
where $S^2_{\hat{\alpha}_{\rm adj}}$ is the bootstrap sample variance for $\hat{\alpha}_{\rm adj}$. Suppose the $\hat{\Delta}_{\rm wadj}$ and $\hat{\Delta}_{\rm IVW}$ are estimators for $\Delta_{\rm wadj}$ and $\Delta_{\rm IVW}$, respectively. Then, we can evaluate the risk-reduction propositions by judging whether $\hat{\Delta}_{\rm est}>0$ for $\text{est} \in \{\text{adj}, \text{wadj}, \text{IVW}\}$, and  the best estimator $\hat{\alpha}_{\rm best}$ is the one with the maximum estimated risk-reduction quantities. Next, we discuss some potential problems of the bootstrap procedures.

In terms of the use of our bootstrap procedure, we caution that the bootstrapping residuals in OLS estimation may not provide valid inference in moderate or high dimension where $p < T_i$ but $p / T_i$ is not close to zero for $i\in \{2, \ldots, n+1\}$ \citep{el2018can}; see alternatives for residual bootstrapping in linear models in \citet{el2018can}.

Recall that $\mathbf{W}^*$  may not be unique if the conditions in Proposition \ref{uniqueness} are not satisfied. Under the setup of our model, non-uniqueness of $\mathbf{W}^*$ would not be a problem for inferential purposes. It is because all risk-reduction propositions and other properties established will still hold. However, non-uniqueness may not be desirable in other model setup. For example, consider the model where $\alpha_i$ is assumed to be not identically distributed, the size of  donor pool to be 2, and there are two solutions to (\ref{SCM}), say, $\mathbf{W}^*_1=(1,0)$ and  $\mathbf{W}^*_2=(0,1)$ with $\var{\alpha_2}\neq \var{\alpha_3}$. In this scenario, the procedure of trying to recover $\alpha_1$ will fail since the resulting $\alpha_1$ will be different with different variances for $\mathbf{W}^*_1=(1,0)$ and  $\mathbf{W}^*_2=(0,1)$. If the non-uniqueness is of concern, users may select the weight that optimizes some objective function. For instance, in the example just discussed, users may select the weight that minimizes the estimated variance of $\hat{\alpha}_{\rm wadj}$.

There are some issues in using $\mc{B}_u$ when $\mathbf{W}^*$ falls in the boundary of the parameter space. See detailed discussions in Section \ref{discussion}.

@@

\begin{table}[H]
  \caption{Bootstrap estimates and results yielded by risk-reduction propositions with $B = 1000$}\label{table2}
  \begin{center}
    \begin{tabular}{lrrr}
      & $\hat{\alpha}_{\rm adj}$ & $\hat{\alpha}_{\rm wadj}$ & $\hat{\alpha}_{\rm IVW}$ \\
      \hline 
    Bootstrapped variance & 0.508 & 0.480 & 2.821 \\
    \end{tabular}
  \end{center}  
\end{table}
\vspace{-.5cm}

Using the parametric bootstrap proposed in Section \ref{varbootstrap}, we estimated parameters for risk-reduction propositions and risk-reduction quantities proposed in  Section \ref{properties}; and the result is presented in Table \ref{table2}. Plugging these estimates into conditions in Section \ref{properties} and risk-reduction formulas in Section \ref{comparisons} yields (1) $\hat{\alpha}_{\text{adj}}$, $\hat{\alpha}_{\text{wadj}}$, and $\hat{\alpha}_{\text{IVW}}$  reduce the risk and (2) the risk-reduction quantities are $28.914$, $28.988$, and $26.444$, respectively. The estimated risk-reduction quantities  vote $\hat{\alpha}_{\rm wadj}$  for the best estimator. Note that this parametric bootstrap procedure is conditioned  on the  donor pool. Then, we verify the consistency of the result yielded by risk-reduction propositions with the reality as below.

@@

@@

Recall that in Section \ref{varbootstrap}, we proposed two possible bootstrap procedures, $\mc{B}_u$  and $\mc{B}_c$. There are some philosophical distinctions between $\mc{B}_u$  and $\mc{B}_c$. $\mc{B}_u$ treats the donor pool as realizations from some infinite super-population of potential donors. In contrast $\mc{B}_c$ treats the donor pool as being fixed  and known before the analysis is conducted, where the randomness comes from parameters and idiosyncratic error.

In this study, for the use of $\mc{B}_u$, we assume $\mathbf{W}^*$ is non-degenerate in the population without taking zero weight.  Recall that in Section \ref{constructionofestimators} we note that if there exists some $\mathbf{W}^*$ satisfies (\ref{SCM}) and $2p < n$, there will be infinitely many solutions to $\mathbf{W}^*$. In the simulation setup, we consider $2p >n$ such that the assumption can be reasonably satisfied. However, in the application, it is likely that $\mathbf{W}^*$ may  take values on the boundary of $\mc{W}$, in which case $\mc{B}_u$ may fail for $\hat{\alpha}_{\mrm{wadj}}$ \citep{andrews2000inconsistency}. For example, if $\mathbf{W}^*$ is observed to fall in the boundary, $\mc{B}_u$ will fail.  Moreover, if $2p < n$, $\mc{B}_u$ fails since the existence of infinitely many solutions is certain (if there exists some $\mathbf{W}^*$ satisfies (\ref{SCM})), and will guarantee degeneracy of  $\mathbf{W}^*$. In this regard, the subsampling in Section 4 of \citet{li2019statistical} may adapt to our scenario to solve this problem by resampling $m$ candidates from donor pool with the conditions $m \to \infty$, and $m/n \to 0$ as $n \to \infty$. But we may leave its mathematical justification for future research. Note that this problem will not occur under  $\mc{B}_c$ because  if $w_i^*=0$ for some $i$ and the weights are known, $w_i^*=0$ makes it impossible for bootstrap estimate of $\hat{\alpha}_i$ to contribute to the bootstrap estimate of distribution of $\hat{\alpha}_{\mrm{wadj}}$.
